---
title: "Parallel programming with R and Azure Batch"
output: 
    html_notebook:
        toc: true
        toc_float: true
        toc_depth: 2
---
<br>
 
# Embarrasingly Parallel Problems
<br>
Also called **perfectly parallel**, these are the most simple cases for parallel computing: 

- Similar tasks are repeated many times
    - Iterations in a loop construct, chunks of data
- Calculations are independent of each other
- Little or no manipulation needed to create parallel tasks
- Examples:  
    - Independent Monte Carlo simulations  
    - Analysis by groups of data  
  
---
Example: Monte Carlo Simulation, random walk of an asset price
```{r}
getClosingPrice <- function(opening_price=100, mean_change=1.001, volatility=0.01, days=1825) { 
    movement <- rnorm(days, mean = mean_change, sd = volatility) 
    path <- cumprod(c(opening_price, movement)) 
    closingPrice <- path[days] 
    return(closingPrice) 
}
```
<br>

To run multiple independent simulations, we can use one of the many looping constructs that come with R, like `for`, `while` or the `apply` family.  


We use `replicate` to run repeatedly in a single thread: 
```{r}
n <- 100000
start_s <- Sys.time() 
closingPrices_s <- replicate(n, getClosingPrice()) 
end_s <- Sys.time()
difftime(end_s, start_s)
hist(closingPrices_s)
```


```{r child = '01_foreach.Rmd'}
```

```{r child = '02_Azure_setup.Rmd'}
```

```{r child = '03_Azure_parallel.Rmd'}
```

# Monitoring and Managing the Azure environment

2.1- Azure file storage
2.2.- Job preparation and release tasks
2.3.- Run multiple tasks in parallel within each compute node. https://docs.microsoft.com/en-us/azure/batch/batch-parallel-node-tasks
2.4.- Job preparation and release tasks. https://docs.microsoft.com/en-us/azure/batch/batch-job-prep-release


- tracking jobs and tasks
- debugging
- checking the environment
- enabling/disabling nodes


```{r child = '04_customize_R_runtime.Rmd'}
```

```{r child = '05_Azure_storage_and_distribution.Rmd'}
```

# References