---
title: "1. foreach"
output: html_notebook
---

# The package `foreach`
## The foreach looping construct
<br>
The package [foreach](https://cran.r-project.org/web/packages/foreach/vignettes/foreach.html) introduces a looping construct that supports parallel execution

The `foreach` loop looks similar to a `for` loop, but is implemented using a binary operator, called `%do%`. Unlike `for`, it returns a value
```{r}
library(foreach)
foreach(i=1:5) %do% sprintf("Hello from iteration %s", i)
```
<br>

The loop `foreach` with the operator `%do%` runs locally and single threaded, pretty much like `for`. This is essentially used for intermediate local tests.

`foreach` comes with another operator, `%dopar%` that runs iterations in parallel. In order to do so, we need to register a parallel backend

## Registering a parallel backend

There are multiple packages that implement functionality to create and register parallel backend clusters:

- `foreach::registerDoSEQ` explicitly register the default sequential backend
- `doParallel::registerDoParallel` local cluster via library(parallel)
- `doFuture::registerDoFuture` HPC with schedulers
- `future::makeClusterMPI` Message Passing Interface (MPI) cluster
- `doAzureParallel::registerDoAzureParallel`

---
Example: local cluster with `doParallel`:
```{r}
library(doParallel)
local_cluster <- parallel::makeCluster(4)
registerDoParallel(local_cluster)

n <- 100000
iterations <- 4
start_p <- Sys.time() 
closingPrices_p <- foreach(i = seq(iterations), .combine = "c") %dopar% {
    replicate(n/iterations, getClosingPrice())
}
end_p <- Sys.time()
difftime(end_p, start_p)
hist(closingPrices_p)
```


## Combining results

`foreach` returns a combination of the result of each iteration. By default, the results are combined into a list as long as the number of iterations. This behaviour can be changed using the parameter `.combine`

In the previous example we specify `.combine = "c"`, hence the results are combined into a vector.

We can as well combine the into columns or rows of a matrix:
```{r}
foreach(i = 1:4, .combine = "cbind") %dopar% rnorm(5)
foreach(i = 1:4, .combine = "rbind") %dopar% rnorm(5)
```

