---
title: "4. R environment in Azure Batch"
output: html_notebook
---

# R runtime environment in Azure Batch

## Installing packages for your parallel runs

```{r echo = FALSE}
# Disable verbose logging
setVerbose(FALSE)
setHttpTraffic(FALSE)
```

Our last call to Azure Batch failed because the package `randomForest` was being used, but was not available on the R runtime environment on Azure Batch. It can be installed on the fly using the `foreach` parameter `.packages`

```{r}
rf <- foreach(ntree = rep(200, 5), .combine = combine, .packages = "randomForest") %dopar%
    randomForest(x, y, ntree = ntree)
rf
```


Note that every time we execute the above chunk, a new installation of the package `randomForest` is triggered.

## Runtime environment of the `doAzureParallel` cluster
<br>
The R code that we ship to run on the Azure Batch cluster runs on a docker container. This ensures a stable runtime environment: all jobs running on the same node run on a new container. The jobs are independent from each other, thus the past jobs history does not affect new jobs, even if previous jobs did things like installing new packages.

A docker image is specified at the time of creating the cluster, in the file `cluster.json`. The default image is "rocker/tidyverse:latest".

## Customizing the docker image for the R runtime

Installing packages at runtime, like in the last example, is not a convenient practice. Recurrent packages should rather be added to the docker image used for all containers. `doAzureParallel` makes this easy by allowing specification of packages in `cluster.json`. Various sources are allowed: cran, github, bioconductor.

It is also possible to specify a custom container image altogether.
